\lstset{ %
language=Python,                % the language of the code
}

For instruction parsing we use Python Lex-Yacc (PLY), an implementation of Lex and Yacc for Python. The Lex part of PLY splits the source in \emph{tokens}, e.g. a \verb!COMMENT!, or a comma. These tokens are passed to Yacc, which finds hierarchical structure in them. When Yacc recognizes a certain pattern, a corresponding action is taken.

\subsection{Lex}

Our lexer splits the source in six tokens: \verb!COMMENT! for comments, \verb!RAW! for compiler directives (lines starting with a dot), \verb!HEX! for hexadecimal values, \verb!INT! for integer values (unsigned), \verb!ID! for label names and instruction names, and \verb!REGISTER! for register identifiers (\verb!$1!, \verb!$2!, etc.)
The following regular expressions specify the tokens.
\begin{quote}
\begin{verbatim}
COMMENT  : #.*
RAW      : \..+
HEX      : 0x[0-9a-f]+
INT      : [0-9]+
ID       : ($L){0,1}[_a-zA-Z0-9][_a-zA-Z0-9\.]*
REGISTER : $[a-z0-9]+
\end{verbatim}
\end{quote}
In addition, we consider the characters `\verb!,!', `\verb!(!', `\verb!)!', `\verb!:!', `\verb!.!', `\verb!+!', and `\verb!-!' as individual tokens, called \emph{literals}.

\subsection{Yacc}
We specified the structure of the assembly language according to the following grammar rules.

\begin{grammar}
<expr> := <raw>
\alt <label>
\alt <comment>
\alt <instr>
\alt <instr> <comment>

<raw> := \verb!RAW!

<label> := \verb!ID! `:'

<comment> := \verb!COMMENT!

<instr> := \verb!ID!
\alt \verb!ID! <arg>
\alt \verb!ID! <arg> `,' <arg>
\alt \verb!ID! <arg> `,' <arg> `,' <arg>

<arg> := <int>
\alt <hex>
\alt <register>
\alt \verb!ID!
\alt \verb!ID! `+' <arg>
\alt \verb!ID! `-' <arg>
\alt <int> `(' <register> `)'
\alt \verb!ID!  `(' <register> `)'

<int> := \verb!INT!
\alt `-' \verb!INT!

<hex> := \verb!HEX!

<register> := \verb!REGISTER!
\end{grammar}

Each expression is stored in a Python object of type \verb!Raw!, \verb!Label!, \verb!Comment!, \verb!Instr! or \verb!Register!.

\subsection{Registers}
To perform dataflow analysis and liveness analysis, we need to parse each 
instruction to determine which registers are set and which registers are used. 
All instructions have a different format, so they need to be parsed seperatly. 
This process is done in the parse_instr module. When we run parse\_instr.parse,
it parses each Instr object from the ir module and sets the register that the
instructoin writes to, the registers that it needs, the offset for writing to 
memory, and immidiate values. An example:
\begin{lstlisting}
if ins.instr == 'add':  
    if len(ins.args) == 3:
        self.gen = [ins.args[0]]
        if self.is_reg(ins.args[2]):
            self.need = [ins.args[1], ins.args[2]]
        else: 
            self.need = [ins.args[1]]
            self.ival = ins.args[2]
    else:
        raise Exception("Invalid number of args for ins: ", ins.instr)   
\end{lstlisting}
